\documentclass[11pt, a4paper]{article}
\usepackage{amsfonts, amsmath, hanging, hyperref, natbib, parskip, times}
\usepackage[pdftex]{graphicx}
\hypersetup{
  colorlinks,
  linkcolor=blue,
  urlcolor=blue
}

\let\section=\subsubsection
\newcommand{\pkg}[1]{{\normalfont\fontseries{b}\selectfont #1}} 
\let\proglang=\textit
\let\code=\texttt 
\renewcommand{\title}[1]{\begin{center}{\bf \LARGE #1}\end{center}}
\newcommand{\affiliations}{\footnotesize}
\newcommand{\keywords}{\paragraph{Keywords:}}

\setlength{\topmargin}{-15mm}
\setlength{\oddsidemargin}{-2mm}
\setlength{\textwidth}{165mm}
\setlength{\textheight}{250mm}

\begin{document}
\pagestyle{empty}

\title{Automatic generation of exams in \proglang{R}}

\begin{center}
  {\bf Bettina Gr\"un$^{1}$, Achim Zeileis$^{2,^\star}$}
\end{center}

\begin{affiliations}
1. Department of Applied Statistics, Johannes Kepler Universit{\"a}t Linz \\[-2pt]
2. Department of Statistics, Universit{\"a}t Innsbruck \\[-2pt]
$^\star$Contact author: \href{mailto:Achim.Zeileis@R-project.org}{Achim.Zeileis@R-project.org}
\end{affiliations}

\keywords Exams, Multiple Choice, Arithmetic Problems

\vskip 0.8cm

Package \pkg{exams} provides a framework for automatic generation of
standardized statistical exams which is especially useful for
large-scale exams. To employ the tools, users just need to supply a
pool of exercises and a master file controlling the layout of the
final PDF document. The exercises are specified in separate Sweave
files (containing \proglang{R} code for data generation and {\LaTeX} code for
problem and solution description) and the master file is a {\LaTeX}
document with some additional control commands. An
overview is given of the main design aims and principles as well as strategies
for adaptation and extension. Hands-on illustrations -- based on
example exercises and control files provided in the package -- are
presented to get new users started easily.


\begin{thebibliography}{}

\bibitem[\protect\citeauthoryear{Gr\"un and Zeileis}{Gr\"un and
  Zeileis}{2009}]{ref1}
Gr\"un, B. and A.~Zeileis (2009).
\newblock Automatic generation of exams in \textsf{R}.
\newblock {\em Journal of Statistical Software\/}~{\em 29\/}(10), 1--14.
\newblock \url{http://www.jstatsoft.org/v29/i10/}

\end{thebibliography}

\end{document}

%% Plain text version

Automatic Generation of Exams in R
Bettina Grün, Achim Zeileis

Package exams provides a framework for automatic generation of
standardized statistical exams which is especially useful for
large-scale exams. To employ the tools, users just need to supply a
pool of exercises and a master file controlling the layout of the
final PDF document. The exercises are specified in separate Sweave
files (containing R code for data generation and LaTeX code for
problem and solution description) and the master file is a LaTeX
document with some additional control commands. This paper gives an
overview of the main design aims and principles as well as strategies
for adaptation and extension. Hands-on illustrations - based on
example exercises and control files provided in the package - are
presented to get new users started easily.

Keywords: exams, multiple choice, arithmetic problems, Sweave, LaTeX, R.

