---
layout: page
#
# Content
#
title: "Written Multiple-Choice Exams with R/exams"
teaser: "Step-by-step guide to generating, conducting, scanning, and automatically evaluating large-scale written exams with exams2nops() in R/exams."
categories:
  - tutorials
tags:
  - exams2nops
  - written
  - mchoice
  - schoice
  - PDF
  - scan
  - evaluation
  - R
author: zeileis

#
# Style
#
image:
  # shown on top of blog post
  title: olympia_front.title.jpg
  # preview in list of posts
  thumb: olympia_front.150.jpg
  # shown on landing page
  # homepage:
  # shown under image on top of blog post
  caption: "R/exams photo (CC-BY)."
---

## Create exam

<div class='row t20 b20'>
  <div class='small-8 medium-9 columns'>

<p>The first step in conducting a <a href="{{ site.url }}/intro/written/">written exam</a> with multiple-choice
(and/or single-choice) exercises in R/exams' NOPS format is to create the exam in PDF format.
First, we load the R <code class="highlighter-rouge">exams</code> package and then simply create a list of exercise file names.</p>

  </div>
  <div class='small-4 medium-3 columns'>
    <img src="{{ site.url }}/images/written-create.svg" alt="create"/>
  </div>
</div>

```{r exam-definition}
library("exams")
myexam <- list(
  "tstat2.Rnw",
  "ttest.Rnw",
  "relfreq.Rnw",
  "anova.Rnw",
  c("boxplots.Rnw", "scatterplot.Rnw"),
  "cholesky.Rnw"
)
```

Here, we use a number of `schoice` and `mchoice` questions that are directly shipped
within the package. In practice, you would use files that you have authored and stored
somewhere locally. Above, exercises in .Rnw format are used but all of the examples
are also available in .Rmd format, leading to virtually identical output.

Then, we create a small exam with only `n = 2` randomly-drawn versions, storing the
resulting PDF files (plus metainformation) on the disk in a new output directory `nops_pdf`.
To customize the exam we assign a different number of points to the different exercises
and also show the respective number of points at the beginning of each question.

```{r exams2nops}
set.seed(403)
ex1 <- exams2nops(myexam, n = 2,
  dir = "nops_pdf", name = "demo", date = "2015-07-29",
  points = c(1, 1, 1, 2, 2, 3), showpoints = TRUE)
```

A random seed is set to make the generated exams exactly reproducible for you
(or ourselves at some point in the future). The output directory now contains three files
that were generated.

```{r exams2nops-files}
dir("nops_pdf")
```

The two PDF files are the two exams we requested above.

`r include_asset("nops_pdf/demo1.pdf")`
`r include_asset("nops_pdf/demo2.pdf")`

Furthermore, the metainfromation about the exam (exam IDs, questions, correct and wrong
answer alternatives) is stored in a `demo.rds` file (serialized R data). This crucial
for being able to evaluate the exam later on.


## Print PDF files

<div class='row t20'>
  <div class='small-8 medium-9 columns'>

<ul>
  <li>A small number of exams can easily be printed on a standard printer.
    Otherwise simply use a print shop.</li>
  <li>It is recommended <em>not</em> to scale the printout (i.e., without "Fit to printable area")
    and to staple the exams in the top-left corner.</li>
  <li>By default the PDFs from <code class="highlighter-rouge">exams2nops()</code> have a blank second page for
    duplex printing (without content on the back of the exam sheet). For non-duplex printing
    simply set <code class="highlighter-rouge">duplex = FALSE</code> when creating the PDFs
    with <code class="highlighter-rouge">exams2nops()</code>.</li>
</ul>

  </div>
  <div class='small-4 medium-3 columns'>
    <img src="{{ site.url }}/images/written-print.svg" alt="print"/>
  </div>
</div>


## Conduct exam

<div class='row t20'>
  <div class='small-8 medium-9 columns'>

<p>The exam is conducted as usual. But if you used the possibilities of
<a href="{{ site.url }}/intro/dynamic/">dynamic exercises</a>
in R/exams, the risk of cheating is greatly reduced.

At the end of the exam you just need to collect the completed exam sheet (first page).
Of course, you can also collect the rest of the exam papers (e.g., to keep future students from
seeing the exercises). However, an advantage of letting the students keep their exercises
reduces the need of having post-exam reviews etc.</p>

  </div>
  <div class='small-4 medium-3 columns'>
    <img src="{{ site.url }}/images/written-exam.svg" alt="exam"/>
  </div>
</div>


## Scan results

<div class='row t20 b20'>
  <div class='small-8 medium-9 columns'>

<p>Each completed exam sheet has information on the exam ID, student ID, and the
checked answers. This needs to be scanned into images (PDF or PNG) and can then
be processed by <code class="highlighter-rouge">nops_scan()</code>.

Typically, it's easy to use the photocopiers provided by your university to scan
all sheets into PDF or PNG files. For example, our university provides us with
Canon ImageRunners and the sheet feeder can easily take about 40-50 sheets and
render them into a single PDF file.</p>

  </div>
  <div class='small-4 medium-3 columns'>
    <img src="{{ site.url }}/images/written-scan.svg" alt="scan"/>
  </div>
</div>

Practical recommendations:
  
* The scanned images become smaller in size if the images are read in just black/white
  (or grayscale). This may sometimes even facilitate extracting the information (see below).
* If the exams were stapled in the top-left corner (see above) the sheet feeder often
  works better if the sheets are rotated by 180 degrees (so that the damaged corner
  is not fed first into the machine). This often improves the scanning results
  considerably and can be accomodated by setting `rotate = TRUE` in `nops_scan()`
  below.

For demonstration, we use two completed demo sheets that are provided along with
the `exams` package and copy them to a dedicated directory `nops_scan`.

```{r exams2nops-scans, results="hide"}
img <- dir(system.file("nops", package = "exams"), pattern = "nops_scan",
  full.names = TRUE)
dir.create("nops_scan")
file.copy(img, to = "nops_scan")
```

`r include_asset("nops_scan/*1.png")`
`r include_asset("nops_scan/*2.png")`

(Note that the first scanned image corresponds to one of the PDFs above while
the other one was generated with custom title/logo/language/etc.)

Using the function `nops_scan()` we can now read all scanned images (i.e.,
all locally available PNG and/or PDF files) and collect everything in a ZIP file.
(Note that if there were PDF files that need to be scanned, then the
PDF toolkit `pdftk` and the function `convert` from ImageMagick need to be
available outside of R on the command line.)

```{r nops_scan, results="hide", warning=FALSE}
nops_scan(dir = "nops_scan")
```

```{r nops_scan-dir}
dir("nops_scan")
```

The resulting file `r include_asset("nops_scan/*.zip")` contains copies of the
PNG files along with a file called `Daten.txt` (for historical reasons) that
contains the scanned information in machine-readable from.

See `?nops_scan` for more details, e.g., multicore support or options for
rotating PDF files prior to scanning.



## Evaluate results

<div class='row t20 b20'>
  <div class='small-8 medium-9 columns'>

<p>In the previous scanning step the exam sheets have just been <em>read</em> but not
yet <em>evaluated</em>, i.e., it has not yet been assessed which questions were answered
(partially) correctly and which were wrong, and no points have been assigned.
Therefore, we use <code class="highlighter-rouge">nops_eval()</code> to carry out these computations and to make
the results available - both in a format easy to read for machines (a CSV file)
and a format for humans (one HTML page for each student).</p>

  </div>
  <div class='small-4 medium-3 columns'>
    <img src="{{ site.url }}/images/written-evaluate.svg" alt="evaluate"/>
  </div>
</div>

To do so, three files are required:

* An RDS file with the exam meta-information, generated by `exams2nops()` above.
* A ZIP file with the scanned sheets, generated by `nops_scan()` above.
* A CSV file (semicolon-separated values) with the student infomation (registration number, name, and some
  ID or username). In practice, this CSV file will typically be processed from some
  registration service or learning management system etc. However, here we simply
  create a suitable CSV file on the fly.

```{r registration-csv}
write.table(data.frame(
  registration = c("1501090", "9901071"),
  name = c("Jane Doe", "Ambi Dexter"),
  id = c("jane_doe", "ambi_dexter")
), file = "Exam-2015-07-29.csv", sep = ";", quote = FALSE, row.names = FALSE)
```

The resulting file is `r include_asset("Exam-*.csv")`.

Now the exam can be evaluated creating an output data frame (also stored
as a CSV file) and individual HTML reports (stored in a ZIP file). Here, we
employ an evaluation scheme without partial points in the multiple-choice questions
and differing points across questions.

```{r nops_eval}
ev1 <- nops_eval(
  register = "Exam-2015-07-29.csv",
  solutions = "nops_pdf/demo.rds",
  scans = Sys.glob("nops_scan/nops_scan_*.zip"),
  eval = exams_eval(partial = FALSE, negative = FALSE),
  interactive = FALSE
)
dir()
```

The evaluated data can be inspected by opening `r include_asset("nops_eval.csv")` in some spreadsheet
software or we can directly look at the data in R. Based on this information, the marks could be
entered into the university's information system.

```{r nops_eval-data}
ev1
```

```{r, echo=FALSE, results="hide"}
unzip("nops_eval.zip")
file.rename(Sys.glob("jane_doe/*.*"), gsub(".html", "-jane_doe.html", Sys.glob("jane_doe/*.*"), fixed = TRUE))
file.rename(Sys.glob("ambi_dexter/*.*"), gsub(".html", "-ambi_dexter.html", Sys.glob("ambi_dexter/*.*"), fixed = TRUE))
```

And `nops_eval.zip` contains subdirectories with HTML reports for each of the two participants.

`r include_asset("jane_doe/*.html")`
`r include_asset("ambi_dexter/*.html")`

These HTML reports could then be uploaded into a learning management system, put on some other
web server, or even sent out via e-mail.
