---
layout: page
#
# Content
#
title: "First Steps"
teaser: "First overview of the structure of R/exams, its exercise templates, and a first hands-on session."
categories:
  - tutorials
tags:
  - Rmd
  - Rnw
  - Markdown
  - LaTeX
  - exams_skeleton
author: zeileis

#
# Style
#
image:
  # shown on top of blog post
  title: laptop_on_table_deriv2.title.jpg
  # preview in list of posts
  thumb: template.150.svg
  # shown on landing page
  # homepage:
  # shown under image on top of blog post
  caption: "R/exams photo (CC-BY)."
---

## Overview

The basic idea of R/exams is to have a [one-for-all exams generator]({{ site.url }}/intro/oneforall/)
where each [exercise]({{ site.url }}/intro/dynamic/) is a standalone file, so that the _same_ exercise can be
rendered easily into a variety output formats such as 
[e-learning systems]({{ site.url }}/intro/elearning/) or
[written exams]({{ site.url }}/intro/written/). To reduce the risk of cheating,
R/exams offers different mechanisms for drawing random variations of exercises:

- Randomly selecting one (or more) out of a set of exercises for each participant.
- Randomly shuffling answer alternatives in single-choice and multiple-choice questions.
- Randomly selecting numbers, text blocks, graphics, etc. using the [R programming language](https://en.wikipedia.org/wiki/R_(programming_language)).

Each exercise (or rather "exercise template") is a standalone file with up to four elements:

1. Data-generating process in R.
2. Question text.
3. Solution text.
4. Meta-information: Type of question (single-choice, multiple-choice, numeric, string, or combinations thereof),
   correct solution, a name/label, etc.

Only the question (2) and the meta-information (4) are mandatory while data
generation (1) and solution (3) are optional. Solutions can be used to either
inform teachers about the correct solution for some randomly drawn exercise - 
or for giving students feedback about how to solve a given exercise. Programming
a good data-generating process is often the hardest part of authoring a new
dynamic exercise but is, of course, crucial for drawing a very large number of
random variations.

For formatting the text within question/solution (e.g., italics, bold face, lists, etc.)
two options are available: [Markdown](https://en.wikipedia.org/wiki/Markdown) or
[LaTeX](https://en.wikipedia.org/wiki/LaTeX). The former is very easy to learn
and recommended for authors who do not know LaTeX yet. If mathematic notation and
formulas are needed, then LaTeX can be used - either within a LaTeX or also within
a Markdown question. The file suffix for [R/Markdown](http://rmarkdown.rstudio.com/)
is `.Rmd` while `.Rnw` is conventionally used for R/LaTeX, known as
[Sweave](http://leisch.userweb.mwn.de/Sweave/).

In the following, a first session should help new R/exams users to explore the materials
available and get a feeling for the workflow. Furthremore, two
[exercise templates]({{ site.url }}/templates/) are shown in more detail:

- [swisscapital]({{ site.url }}/templates/swisscapital/): A static single-choice 
  knowledge quiz question where the only randomization is a subsampling of the presented
  alternatives.
- [deriv]({{ site.url }}/templates/deriv/): A dynamic airthmetic question about the
  product rule for derivatives where various numbers are drawn randomly in R.



## A first session

To explore what is available in R/exams two things are necessary: a couple of exercise
templates to draw inspiration from, and some demo R scripts that illustrate how the
different `exams2xyz()` functions can produce output in a variety of formats. (Here,
`xyz` is a _placeholder_ for the output format like `html`, `pdf`, `moodle`, etc.)
The function `exams_skeleton()` is a helper function that generates exactly these
materials in a specified folder. For example, when `exams` has already been
[installed]({{ site.url }}/tutorials/installation/), the following can be carried out
in R:

```{r skeleton, results="hide"}
library("exams")
exams_skeleton(markup = "markdown", encoding = "UTF-8",
  writer = c("exams2html", "exams2pdf", "exams2moodle"))
```

This copies all R/Markdown files to the current working directory along with demo scripts,
illustrating the usage of `exams2html()` and `exams2pdf()` for customizable HTML and
PDF output, respectively, along with Moodle output via `exams2moodle()`. Specifically,
the working directory then contains: 

```{r output}
dir()
```

Simply open the demo script 
`r include_asset("demo-all.R")` for the first steps and then continue with
more details in `r include_asset("demo-html.R")`, `r include_asset("demo-pdf.R")`, or
`r include_asset("demo-moodle.R")`, respectively.
More information about all the `exercises` can be found in the
[exercise template gallery]({{ site.url }}/templates/) online.


## Single-choice: Swiss capital

Knowledge quiz question (about the Swiss capital) with 1 correct and 6 false alternative.
4 out of the 6 false alternatives are sampled randomly and shuffled. Otherwise the
exercise is static and contains no R code. Below the structure of the exercise is
highlighted side-by-side in both `.Rmd` and `.Rnw`, illustrating that the additional
exercise markup is similar in spirit to the respective typesetting system.

To see what the output looks like, run the following code within R or take a look at
[swisscapital]({{ site.url }}/templates/swisscapital/) in the exercise gallery.


<div class='row t20 b20'>
  <div class='small-5 medium-3 large-2 columns'>
    <b>Markdown</b><br/>
    <img src="{{ site.url }}/images/template-rmd.svg" alt="Rmd" />
  </div>
  <div class='small-1 medium-3 large-4 columns'>
  <p></p>
  </div>
  <div class='small-5 medium-3 large-2 columns'>
    <b>LaTeX</b><br/>
    <img src="{{ site.url }}/images/template-rnw.svg" alt="Rnw" />
  </div>
  <div class='small-1 medium-3 large-4 columns'>
  <p></p>
  </div>
</div>

<div class='row t20 b20'>
  <div class='medium-6 columns'>
    <pre>
library("exams")
exams2html("swisscapital.Rmd")
exams2pdf("swisscapital.Rmd")
    </pre>
  </div>
  <div class='medium-6 columns'>
    <pre>
library("exams")
exams2html("swisscapital.Rnw")
exams2pdf("swisscapital.Rnw")
    </pre>
  </div>
</div>

#### Question

The question is straightforward and lists 7 answer alternatives.
The correct solution (Bern) is declared in the meta-information below.

<div class='row t20 b20'>
  <div class='medium-6 columns'>
    <b>Markdown</b>
    <pre>
Question
========
What is the seat of the federal
authorities in Switzerland
(i.e., the de facto capital)?

Answerlist
----------
* Basel
* Bern
* Geneva
* Lausanne
* Zurich
* St. Gallen
* Vaduz
    </pre>
  </div>
  <div class='medium-6 columns'>
    <b>LaTeX</b>
    <pre>

\begin{question}
What is the seat of the federal
authorities in Switzerland
(i.e., the de facto capital)?

\begin{answerlist}
  \item Basel
  \item Bern
  \item Geneva
  \item Lausanne
  \item Zurich
  \item St.~Gallen
  \item Vaduz
\end{answerlist}
\end{question}
    </pre>
  </div>
</div>


#### Solution

The optional solution can provide some general feedback (that explains
why Bern is correct in this question) as well as more details for each
answer alternative (only listing true/false here). It is also possible 
to have _no solution_, or _just a general text_, or _just the answer list_.

<div class='row t20 b20'>
  <div class='medium-6 columns'>
    <b>Markdown</b>
    <pre>
Solution
========
There is no de jure capital but
the de facto capital and seat of
the federal authorities is Bern.

Answerlist
----------
* False
* True
* False
* False
* False
* False
* False
    </pre>
  </div>
  <div class='medium-6 columns'>
    <b>LaTeX</b>
    <pre>

\begin{solution}
There is no de jure capital but
the de facto capital and seat of
the federal authorities is Bern.

\begin{answerlist}
  \item False.
  \item True.
  \item False.
  \item False.
  \item False.
  \item False.
  \item False.
\end{answerlist}
\end{solution}
    </pre>
  </div>
</div>


#### Meta-information

In the meta-information, each exercise must specify at least an `extype`,
and an `exsolution`, and should typically also have a short descriptive
`exname`. The type for single-choice questions is `schoice` and the
`exsolution` is simply a binary coding of the 7 answer alternatives.
For randomization, additionally `exshuffle` is set to 5 so that 1 correct
and 4 random wrong alternatives are subsampled and shuffled.


<div class='row t20 b20'>
  <div class='medium-6 columns'>
    <b>Markdown</b>
    <pre>
Meta-information
================
exname: Swiss Capital
extype: schoice
exsolution: 0100000
exshuffle: 5
    </pre>
  </div>
  <div class='medium-6 columns'>
    <b>LaTeX</b>
    <pre>


\exname{Swiss Capital}
\extype{schoice}
\exsolution{0100000}
\exshuffle{5}
    </pre>
  </div>
</div>


## Numeric: Product rule for derivatives

Arithmetic question for computing the first derivative of a product function with
two factors, using the product rule. The parameters of the function are drawn
randomly in R and subsequently inserted into the question and the solution (which
is not shown below for brevity). LaTeX is used for typesetting the mathematical
formulas - both in the Markdown and the LaTeX version.

To see what the output looks like, run the following code within R or take a look at
[deriv]({{ site.url }}/templates/deriv/) in the exercise gallery.


<div class='row t20 b20'>
  <div class='small-5 medium-3 large-2 columns'>
    <b>R/Markdown</b><br/>
    <img src="{{ site.url }}/images/template-rmd.svg" alt="Rmd" />
  </div>
  <div class='small-1 medium-3 large-4 columns'>
  <p></p>
  </div>
  <div class='small-5 medium-3 large-2 columns'>
    <b>R/LaTeX</b><br/>
    <img src="{{ site.url }}/images/template-rnw.svg" alt="Rnw" />
  </div>
  <div class='small-1 medium-3 large-4 columns'>
  <p></p>
  </div>
</div>

<div class='row t20 b20'>
  <div class='medium-6 columns'>
    <pre>
library("exams")
exams2html("deriv.Rmd")
exams2pdf("deriv.Rmd")
    </pre>
  </div>
  <div class='medium-6 columns'>
    <pre>
library("exams")
exams2html("deriv.Rnw")
exams2pdf("deriv.Rnw")
    </pre>
  </div>
</div>

#### Data generation

The R code simply draws the two parameters `a` and `b` as well as the argument `c`
from a discrete uniform distribution. (See the question below for the actual function used.)
Subsequently, the random numbers are inserted into the correct solution and storing
it as `res`.

<div class='row t20 b20'>
  <div class='medium-6 columns'>
    <b>R/Markdown</b>
    <pre>
```{r, echo=FALSE, results="hide"}
## parameters
a <- sample(2:9, 1)
b <- sample(seq(2, 4, 0.1), 1)
c <- sample(seq(0.5, 0.8, 0.01), 1)
## solution
res <- exp(b * c) * (a * c^(a-1) +
       b * c^a)
```
    </pre>
  </div>
  <div class='medium-6 columns'>
    <b>R/LaTeX</b>
    <pre>
<<echo=FALSE, results=hide>>=
## parameters
a <- sample(2:9, 1)
b <- sample(seq(2, 4, 0.1), 1)
c <- sample(seq(0.5, 0.8, 0.01), 1)
## solution
res <- exp(b * c) * (a * c^(a-1) +
       b * c^a)
@
    </pre>
  </div>
</div>


#### Question

The question asks for the derivative of a particular function and simply inserts
the parameters drawn above into the question template. (The solution is set up
analogously in the exercise template but not shown here for brevity.)

<div class='row t20 b20'>
  <div class='medium-6 columns'>
    <b>R/Markdown</b>
    <pre>
Question
========
What is the derivative of
$f(x) = x^{`r a`} e^{`r b` x}$,
evaluated at $x = `r c`$?
    </pre>
  </div>
  <div class='medium-6 columns'>
    <b>R/LaTeX</b>
    <pre>
\begin{question}
What is the derivative of
$f(x) = x^{\Sexpr{a}} e^{\Sexpr{b}x}$,
evaluated at $x = \Sexpr{c}$?
\end{question}
    </pre>
  </div>
</div>


#### Meta-information

Again, the meta-information lists `extype` (which is `num` for numeric exercises),
`exsolution` (which is inserted dynamically from the variable `res` computed above),
and a short `exname`. Additionally, the tolerance `extol` which is used for automatically
evaluating answers, e.g., in an online test/exam.

<div class='row t20 b20'>
  <div class='medium-6 columns'>
    <b>R/Markdown</b>
    <pre>
Meta-information
================
extype: num
exsolution: `r fmt(res)`
exname: derivative exp
extol: 0.01
    </pre>
  </div>
  <div class='medium-6 columns'>
    <b>R/LaTeX</b>
    <pre>


\extype{num}
\exsolution{\Sexpr{fmt(res)}}
\exname{derivative exp}
\extol{0.01}
    </pre>
  </div>
</div>
