---
title: "R/exams questions"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include = FALSE}
library("exams2learnr")
```

## Knowledge quiz (single-choice)

```{r schoice1, echo = FALSE, message = FALSE}
exams2learnr("swisscapital.Rmd", allow_retry = TRUE, incorrect = "Incorrect, try again.")
```

## Knowledge quiz (multiple-choice)

```{r mchoice2, echo = FALSE, message = FALSE}
exams2learnr("capitals.Rmd")
```

## Arithmetic (numeric)

_Important note:_ Numeric questions are currently handled like string questions.
Thus, only a `.` (and not a `,`) will work as the decimal separator
and the exact digits have to be supplied. No tolerance is applied
and missing or extra trailing zeros etc.

```{r num3, echo = FALSE, message = FALSE}
exams2learnr("deriv.Rnw", allow_retry = TRUE)
```

## Arithmetic (single-choice)

```{r schoice4, echo = FALSE, message = FALSE}
exams2learnr("deriv2.Rnw")
```

## Multiple-choice with graphic

```{r mchoice5, echo = FALSE, message = FALSE}
exams2learnr("boxplots.Rmd")
```

## Multiple-choice with R output

```{r mchoice6, echo = FALSE, message = FALSE}
exams2learnr("ttest.Rmd")
```

## String question

```{r mchoice7, echo = FALSE, message = FALSE}
exams2learnr("function.Rmd", allow_retry = TRUE)
```
