---
layout: page
#
# Content
#
title: "tstat"
teaser: "tstat exercise."
categories:
  - templates
tags:
  - num
  - arithmetic
  - t statistic
  - statistics
author: zeileis

---

```{r tstat-files, echo=FALSE, results="hide", warning=FALSE}
library("exams")
ex <- system.file("exercises", c("tstat.Rnw", "tstat.Rmd"), package = "exams")
file.copy(ex, to = getwd())
set.seed(1090)
exams2html("tstat.Rnw", name = "tstat-Rnw-html-", dir = ".")
set.seed(1090)
exams2html("tstat.Rmd", name = "tstat-Rmd-html-", dir = ".")
set.seed(1090)
exams2pdf("tstat.Rnw", name = "tstat-Rnw-pdf-", dir = ".")
set.seed(1090)
exams2pdf("tstat.Rmd", name = "tstat-Rmd-pdf-", dir = ".")
set.seed(1090)
Sweave("tstat.Rnw")
set.seed(1090)
knitr::knit("tstat.Rmd", quiet = TRUE, encoding = "UTF-8")
```

**Exercise templates:** `r include_asset("tstat.Rnw")` \| `r include_asset("tstat.Rmd")`

**Exercise weaved:** `r include_asset("tstat.tex")` \| `r include_asset("tstat.md")`

**PDF output:**

`r include_asset("tstat-Rnw-pdf-1.pdf")`
`r include_asset("tstat-Rmd-pdf-1.pdf")`

**HTML output:**

`r include_asset("tstat-Rnw-html-1.html")`
`r include_asset("tstat-Rmd-html-1.html")`

_(Note that the HTML output contains mathematical equations in MathML.
It is displayed by browsers with MathML support like Firefox or Safari - but not Chrome.)_

**Demo code:**

```{r tstat-demo, eval=FALSE}
library("exams")

set.seed(1090)
exams2html("tstat.Rnw")
set.seed(1090)
exams2pdf("tstat.Rnw")

set.seed(1090)
exams2html("tstat.Rmd")
set.seed(1090)
exams2pdf("tstat.Rmd")
```

